Effects on projectivity ratings by Embedding Operator and Trigger --- \newline Data Analysis

### 2022-11-28 materials-lisa/analysis/3_projection-across-operators/analysis.Rnw
updates to analysis and added abstract file

  \item We do not see any group of verbs that could be characterized as \lq semi-factive\rq\ in the sense of Karttunen.
  \begin{itemize}
    \item Specifically, \emph{discover} does not follow the predicted pattern: It is not more projective under negation, but most projective in conditionals and questions.
    \item For Karttunnen's \lq factives\rq, no difference between operators is expected.
    \item Kajsa Djärv about this distinction: cognitive predicates are semi-factive, and emotives are factive. The pattern suggested by karrtunen is also not found here
  \end{itemize}

@@

@
Some more generalizations: We find different \lq profiles\rq for different verbs, of how embedding operators affect projectivity based on the verb. Groups of verbs show similar profiles:

@@

% \section{Analysis}
% 
% We are interested in how the response (projectivity ratings) depends on the verb and embedding operator, and whether the above generalizations can be supported statistically.
% 
% 
% 
% <<>>=
% formula <- response ~ op * verb
% # formula <- response ~ op * verb + (1 | workerid) + (1 | item)
% # formula <- response ~ op * verb + (1 + op + verb | workerid) + (1 + op + verb | item)
% @
% 
% Trying a beta-regression. Our response variable has observations from a sliding scale between $(0,1)$. For beta regression, we rescale it to $[0,1]$, using method used in Degen \& Tonhauser (2022), from Smithson \& Verkuilen (2006).
% 
% <<>>=
% data$response = (data$projective*(nrow(data)-1) + .5)/nrow(data)
% @
% 
% 
% <<>>=
% # library(rstan)
% # options(mc.cores=parallel::detectCores())
% # rstan_options(auto_write=TRUE)
% # library(brms)
% # 
% # model1 <- brm(formula = bf(response ~ op * verb ,
% #                            phi ~ op * verb,
% #                            family = Beta()),
% #               family = Beta(),
% #               data = data,
% #               iter=10000, warmup=2000,chains=6)
% 
% 
% @
% 
%

>>>>

  \item We do not see any group of verbs that could be characterized as \lq semi-factive\rq\ in the sense of Karttunen.
  \begin{itemize}
    \item Specifically, \emph{discover} does not follow the predicted pattern: It is not more projective under negation, but most projective in conditionals and questions.
    \item For Karttunnen's \lq factives\rq, no difference between operators is expected.
    \item Kajsa Djärv about this distinction: cognitive predicates are semi-factive, and emotives are factive. The pattern suggested by karrtunen is also not found here (if anything, it might be the other way round)
  \end{itemize}

@@

@
Some more generalizations: We find different \lq profiles\rq\ for different verbs, of how embedding operators affect projectivity based on the verb. Groups of verbs show similar profiles:

@@

\section{Analysis}

We are interested in how the response (projectivity ratings) depends on the verb and embedding operator, and whether the above generalizations can be supported statistically.\\

% \subsection{Frequentist modeling}
Exploring a couple of models

\subsection{Linear frequentist with \texttt{suggest/m} as baseline}

<<freq-linear-suggest-m, message=FALSE, warning=FALSE>>=

# our independent variables
## inference-triggering predicate, coded as "verb"
data$verb <- as.factor(data$verb)
levels(data$verb)
# contrasts(data$verb)
### dummy coding with "suggest" as baseline
data$verb <- relevel(data$verb, ref = "suggest")
# contrasts(data$verb)

## embedding operator, coded as "op"
data$op <- as.factor(data$op)
levels(data$op)
# contrasts(data$op)
### dummy coding with "m" as baseline
data$op <- relevel(data$op, ref = "m")
# contrasts(data$op)

# coding random effects as factors
data$workerid <- as.factor(data$workerid)
data$content <- as.factor(data$content)

library(lme4)
library(lmerTest)

# glmm1 <- lmer(projective ~ verb * op + (1 | workerid) +
#                  (1 | content), data=data)

# save.image("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
load("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
print(summary(glmm1), cor=F, dig=3)
@

\subsection{Linear frequentist with \texttt{annoyed/q} as baseline}
<<freq-linear-annoyed-q, message=FALSE>>=

# dummy coding for "verb" with "suggest" as baseline
data$verb <- relevel(data$verb, ref = "be_annoyed")

### dummy coding for "op" with "q" as baseline
data$op <- relevel(data$op, ref = "q")

glmm2a <- lmer(projective ~ verb * op + (1 | workerid) +
                 (1 | content), data=data)

# save.image("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
load("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
print(summary(glmm2a), cor=F, dig=3)

\subsection{Linear frequentist with \texttt{prove/n} as baseline}
<<freq-linear-annoyed-q, message=FALSE>>=

# dummy coding for "verb" with "prove" as baseline
data$verb <- relevel(data$verb, ref = "prove")

### dummy coding for "op" with "n" as baseline
data$op <- relevel(data$op, ref = "n")

# glmm3 <- lmer(projective ~ verb * op + (1 | workerid) +
#                  (1 | content), data=data)
# 
# save.image("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
load("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
print(summary(glmm3), cor=F, dig=3)

\subsection{Linear frequentist with \texttt{know/m} as baseline}
<<freq-linear-annoyed-q, message=FALSE>>=

# dummy coding for "verb" with "prove" as baseline
data$verb <- relevel(data$verb, ref = "know")

### dummy coding for "op" with "n" as baseline
data$op <- relevel(data$op, ref = "m")

# glmm4 <- lmer(projective ~ verb * op + (1 | workerid) +
#                  (1 | content), data=data)
# 
# save.image("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
load("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
print(summary(glmm4), cor=F, dig=3)

% \subsection{Bayesian beta-regression with \texttt{suggest/m} as baseline}
% 
% Our response variable has observations from a sliding scale between $(0,1)$. For beta regression, we rescale it to $[0,1]$, using method used in Degen \& Tonhauser (2022), from Smithson \& Verkuilen (2006).
% 
% <<transform-for-beta>>=
% data$response = (data$projective*(nrow(data)-1) + .5)/nrow(data)
% @
% 
% <<betareg>>=
% library(rstan)
% options(mc.cores=parallel::detectCores())
% rstan_options(auto_write=TRUE)
% library(brms)
% 
% 
% 
% # bbmm1 <- brm(formula = bf(response ~ op * verb, # beta model mean parameter (mu)
% #                             phi ~ op * verb, # beta model precision parameter (phi)
% #                             family = Beta()),
% #                family = Beta(),
% #                data = data,
% #                iter=10000, warmup=2000,chains=6)
% 
% @

% \subsubsection{Linear frequentist with difference coding}
% Make sure I understand how difference coding works
% This should help
% https://stats.oarc.ucla.edu/r/library/r-library-contrast-coding-systems-for-categorical-variables/#backward

% https://rdrr.io/github/tsostarics/contrastable/src/R/forward_difference_code.R

% <<>>=
% # frequentist fixed effects only, input error: no matrix
% # bm1 <- betareg(
% #   response ~ verb * op | # beta model mean parameter (mu)
% #     response ~ verb * op, # beta model precision parameter
% #   data = data,
% #   link = "logit"
% # )
% formula <- response ~ op * verb
% # formula <- response ~ op * verb + (1 | workerid) + (1 | item)
% # formula <- response ~ op * verb + (1 + op + verb | workerid) + (1 + op + verb | item)
% @
% 
% Trying a beta-regression. 
% <<>>=
% # library(rstan)
% # options(mc.cores=parallel::detectCores())
% # rstan_options(auto_write=TRUE)
% # library(brms)
% # 
% # model1 <- brm(formula = bf(response ~ op * verb ,
% #                            phi ~ op * verb,
% #                            family = Beta()),
% #               family = Beta(),
% #               data = data,
% #               iter=10000, warmup=2000,chains=6)
% 
% 
% @
% 
%
